### 2019-01-29 paper_epjc.tex

%\RequirePackage{fix-cm}

\documentclass[pdftex,twocolumn,epjc3]{svjour3}          % twocolumn

\RequirePackage[T1]{fontenc}

\smartqed  % flush right qed marks, e.g. at end of proof

\RequirePackage{graphicx}
\RequirePackage{mathptmx}      % use Times fonts if available on your TeX system
\RequirePackage{flushend}
\RequirePackage[numbers,sort&compress]{natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}

\journalname{Eur. Phys. J. C}

\begin{document}

\title{Insert your title here\thanksref{t1}}

\subtitle{Do you have a subtitle?\\ If so, write it here}

\author{First Author\thanksref{e1,addr1}
        \and
        Second Author\thanksref{e2,addr2,addr3} %etc.
}

\thankstext[$\star$]{t1}{Thanks to the title}
\thankstext{e1}{e-mail: magic1@xxx.xx}
\thankstext{e2}{e-mail: magic2@xxx.xx}

\institute{First Address, Street, City, Country\label{addr1}
          \and
          Second Address, Street, City, Country\label{addr2}
          \and
          \emph{Present Address:} Street, City, Country\label{addr3}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Insert your abstract here. Insert your abstract here. Insert your abstract here.
Insert your abstract here. Insert your abstract here. Insert your abstract here.
Insert your abstract here. Insert your abstract here. Insert your abstract here.
\end{abstract}

\begin{thebibliography}{9}
\bibliography{paper}
\end{thebibliography}

\end{document}