### 2019-01-31 paper.tex
Updates from ShareLaTeX

\documentclass[a4paper,11pt]{article}
\pdfoutput=1 % if your are submitting a pdflatex (i.e. if you have
             % images in pdf, png or jpg format)

\usepackage{jheppub} % for details on the use of the package, please
                     % see the JHEP-author-manual

\usepackage[T1]{fontenc} % if needed
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
%\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{\boldmath Heavy resonances at energy frontier colliders}

\author[a,c]{David~Jamin}
\author[a]{\!\!, Clement~Helsens}
\author[a]{\!\!, Michelangelo~L.~Mangano}
\author[b]{\!\!, Thomas~G.~Rizzo}
\author[a]{and Michele~Selvaggi}

\affiliation[a]{CERN, CH-1211 Geneva 23, Switzerland}
\affiliation[b]{SLAC National Accelerator Laboratory 2575 Sand Hill Rd., Menlo Park, CA, 94025 USA}
\affiliation[c]{Academia Sinica, Institute of  Physics, Taipei, Taiwan}

% e-mail addresses: one for each author, in the same order as the authors
\emailAdd{david.jamin@cern.ch}
\emailAdd{clement.helsens@cern.ch}
\emailAdd{michelangelo.mangano@cern.ch}
\emailAdd{rizzo@slac.stanford.edu}
\emailAdd{michele.selvaggi@cern.ch}

\newcommand{\CH}[1] {\textbf{\textcolor{blue}{CH - #1}}}
\newcommand{\MS}[1] {\textbf{\textcolor{purple}{MS - #1}}}
\newcommand{\MLM}[1] {\textbf{\textcolor{green}{MLM - #1}}}
\newcommand{\TR}[1] {\textbf{\textcolor{red}{TR - #1}}}
\newcommand{\DA}[1] {\textbf{\textcolor{orange}{DJ - #1}}}

\newcommand{\Zp}{\ensuremath{Z^{\prime}}}
\newcommand{\ZpSSM}{\ensuremath{Z^{\prime}_{\mathrm{SSM}}}}
\newcommand{\Z}{\ensuremath{Z}}
\newcommand*{\Zptata}{\ensuremath{Z^{\prime}\rightarrow \tau\tau}}
\newcommand*{\Zpee}{\ensuremath{Z^{\prime}\rightarrow \text{ee}}}
\newcommand*{\Zpmumu}{\ensuremath{Z^{\prime}\rightarrow \mu\mu}}
\newcommand*{\Zpll}{\ensuremath{Z^{\prime}\rightarrow \ell\ell}}
\newcommand*{\Zptt}{\ensuremath{Z^{\prime} \rightarrow \ttbar}}
\newcommand{\ptSub}[1]{\ensuremath{p_{\text{T} #1}}}
\newcommand{\ptSup}[1]{\ensuremath{p_{\text{T}}^{#1}}}
\newcommand{\pt}{\ensuremath{p_{\text{T}}}}
\newcommand{\ptEl}{\ensuremath{p_{\text{T}}^{e}}}
\newcommand{\ptMu}{\ensuremath{p_{\text{T}}^{\mu{}}}}
\newcommand{\ptZp}{\ensuremath{p_{\text{T}}^{\ensuremath{Z^{\prime}}}}}
\newcommand{\mZp}{\ensuremath{M_{\ensuremath{Z^{\prime}}}}}
\newcommand{\mSD}{\ensuremath{m_{\ensuremath{SD}}}}
\newcommand{\mll}{\ensuremath{M_{\ensuremath{ll}}}}
\newcommand*{\hht}{\ensuremath{H_{\ensuremath{T}}}}
\newcommand*{\met}{\ensuremath{E_{T}^{\rm{miss}}}}
\newcommand*{\intlumifcc}{\ensuremath{\mathcal{L}=30\,\text{ab}^{-1}}}
\newcommand*{\intlumihelhc}{\ensuremath{\mathcal{L}=15\,\text{ab}^{-1}}}
\newcommand*{\qjj}{\ensuremath{Q^{*} \rightarrow jj}}
\newcommand*{\rsg}{\ensuremath{G_{RS} \rightarrow W^+W^-}}
\newcommand*{\vj}{\ensuremath{\text{V + jets}}}
\newcommand*{\metvec}{\vec{\pt}^{\textrm{miss}}}

\newcommand*{\ee}{\ensuremath{e^{+}e^{-}}}
\newcommand*{\mumu}{\ensuremath{\mu^{+}\mu^{-}}}
\newcommand*{\tautau}{\ensuremath{\tau^{+}\tau^{-}}}
\newcommand*{\ellell}{\ensuremath{\ell^{+}\ell^{-}}}
\newcommand*{\ttbar}{\ensuremath{t\bar{t}}}
\newcommand*{\bbbar}{\ensuremath{b\bar{b}}}
\newcommand*{\ww}{\ensuremath{W^{+}W^{-}}}
\newcommand*{\jj}{\ensuremath{jj}}
\newcommand*{\com}{c.o.m.~}
\newcommand*{\cl}{C.L.~}

\newcommand{\amc}{{\sc MadGraph5\textunderscore}a{\sc MC@NLO}}
\newcommand*{\MGAMC}{\textsc{MG5}\_aMC}

\newcommand{\py}{{\sc Pythia8}}
\newcommand{\delphes}{{\sc Delphes3}}
\newcommand*{\pdf}[1]{\texttt{#1}}

\def\ifb{\mbox{fb$^{-1}$}}%  Inverse femtobarns.
\def\afb{\mbox{ab$^{-1}$}}%  Inverse atobarns.

\def\TeV{\ifmmode {\mathrm{\ Te\kern -0.1em V}}\else
                   \textrm{Te\kern -0.1em V}\fi}%
\def\GeV{\ifmmode {\mathrm{\ Ge\kern -0.1em V}}\else
                   \textrm{Ge\kern -0.1em V}\fi}%
\def\MeV{\ifmmode {\mathrm{\ Me\kern -0.1em V}}\else
                   \textrm{Me\kern -0.1em V}\fi}%
\def\keV{\ifmmode {\mathrm{\ ke\kern -0.1em V}}\else
                   \textrm{ke\kern -0.1em V}\fi}%
\def\eV{\ifmmode  {\mathrm{\ e\kern -0.1em V}}\else
                   \textrm{e\kern -0.1em V}\fi}%

\abstract{This paper explores the physics reach of the proton-proton Future Circular Collider (FCC-hh) and High Energy LHC (HE-LHC) for searches of new particles decaying to two high energetic leptons, jets (non-tops), tops and W/Z boson. We discuss the expected exclusion limits and discovery potential for benchmark models predicting new massive particles that result in resonant structures in the invariant mass spectrum. This document also includes a detailed study on the discrimination potential of a $\Zp$ at HE-LHC in case of discovery at the end of High-Luminosity LHC (HL-LHC).}

\begin{document}
\maketitle
\flushbottom

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
Particle accelerators are built to answer some of the most fundamental questions about the natural world. For LHC it was guaranteed that the Standard Model (SM) Higgs boson would be found; however, for the next generation of machines there is no possibility to guarantee that any new particle will be discovered. Still with much higher center of mass energies compared to LHC, there are guaranteed deliverables like the study of the Higgs and top-quark properties and exploration of electroweak symmetry breaking phenomena with unmatchable precision and sensitivity.

New machines are built to make direct discoveries, and even though we do not have any guaranted discoveries, we need to make sure that we will cover a large fraction of Beyond the SM (BSM) phase space. Qualitatively, the mass reach should roughly scale by a factor of $\sqrt{s}/14$, so $\sim 7$ for $\sqrt{s}=100$. However, with the parton luminosities evolving with $Q^2$, thus dropping a bit faster at high energy, one typically finds it is only a factor of $\sim 5$ increase. But the statistics are also enhanced by several orders of magnitude for many BSM phenomena, that the LHC could barely touch during its exploitation. It is not only that the mass reach increases by a large factor, but it is the fact that if the LHC were to see some hints of possible new physics, by increasing the energy by a factor seven, we would increase the statistics by two or even three orders of magnitude, and we can use this new machine to study with great accuracy what it is that has been found. In addition we could have the ability to provide firm answers to questions such as: is the SM dynamics all there at the TeV scale, is there a TeV scale solution to the hierarchy problem, is dark matter a thermal wimp (either we discover it as a WIMP, or we discover DM is not a WIMP and it has to be something else), was the cosmological EW phase transition 1st order, etc...

Concerning the topic of this paper, the discovery potential for new resonances not predicted by the SM makes Future Circular Colliders (FCC) the only place to search for such heavy particles, compared to the current LHC and the coming HL-LHC.
In the framework of FCC it is also extremely relevant to discuss the main limitations of the detector in identifying high energetic top-quarks or W/Z bosons. Indeed 100\,TeV proton-proton collisions will produce very large quantities of multi-TeV's bosons or top, thus the design of the detector needs detailed optimisations in order to achieve the required physics goals. The capabilities of such a detector should include the capabilities of measuring multi-TeV leptons, top-quarks and bosons, and will be discussed in this paper.

This document presents the expectations of some of the most relevant BSM scenarios and the considered models are discussed in the section~\ref{sec:bsmmodels}.
Details on the generators, detector parameterization, statistical methods and other analysis techniques developed are presented in the section~\ref{sec:simulation}.
The leptonic resonances ($ee$, $\mu\mu$, $\tau\tau$) and the hadronic resonances (WW, $t\bar{t}$ and jj) analyses are detailed in sections~\ref{sec:lep} and~\ref{sec:hadronic} respectively.
The analyses strategy started from optimisation at FCC-hh, but the results have also been extracted in the case of the HE-LHC scenario and are summarised in section~\ref{sec:ana27tev}.
Finally, a study of model discrimination at HE-LHC in case of discovery at the end of HL-LHC run is presented in section~\ref{sec:zprimedisc}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{BSM Models}
\label{sec:bsmmodels}

In order to explore and contrast the capabilities of future colliders to discover and examine the properties of new physics, a broad set of benchmark models needs to be employed. In the
case of new heavy resonances, this benchmark set should be sufficiently complete such that all of the major discovery channels of relevance are represented. As discussed above, here we
are particularly interested in the 2-body final states of these resonances (since they are generally dominant in almost all new physics scenarios) consisting of opposite sign dilepton
pairs ($e^+e^-, \mu^+\mu^-$ and $\tau^+\tau^-$), dijets, $ t\bar t$ and $W^+W^-$.  We note that it is highly likely that at least one or possibly more of these 2-body channels will posses
a respectable branching fraction for the new resonances that result from any specific beyond the Standard Model scenario. Note that decays into pairs of secondary objects that
then themselves decay hadronically can often populate the dijet channel if the final state jets are sufficiently boosted so this channel can represent many different final states unless
substructure studies are performed.  When there are 2 or more of these channels available for simultaneous study we have an increased chance of learning significantly more about
the underlying physics model behind the new resonance. The most important properties of a newly discovered resonance that need to be determined (other than the mass) are its
production cross section, which will sometimes require a good understanding of the underlying background shape especially for a broad resonance and its spin (as was the case in
the example of the Higgs boson). These properties alone can provide important information about the BSM model from which the signal originated. The spin measurement usually requires
the reconstruction of the angular distribution of the resonance decay products and, hence, a respectable amount of statistics although the observation of certain final states can
immediately exclude some spin possibilities as was the case with $H\rightarrow \gamma \gamma$.

A new, neutral, spin-1 gauge boson, $\Zp$, which is usually a color-singlet object produced in the $q\bar q$ channel, is a ubiquitous feature of many models that predict new
physics~\cite{Langacker:2008yv,Rizzo:2006nw,Carena:2004xs,Salvioni:2009mt}.  While falling into several distinct classes, $\Zp$ are most commonly associated with the
extension of the the SM electroweak gauge group by an additional U(1) or SU(2) factor although more significant augmentations are possible. When the additional factor is non-abelian,
as in the case of SU(2), a new $W^\pm$' gauge boson generally also appears in the spectrum together with the $\Zp$ and with a comparable mass.  Of this subset of models, those
that arise from Grand Unified Theory frameworks are the ones most commonly encountered in the literature and include familiar examples such as the Left-Right Symmetric
Model (LRM)~\cite{Senjanovic:1975rk,Senjanovic:1975rk,Mohapatra:1980yp} which results from SO(10)
(or larger GUT groups) and where the SM is augmented by an SU(2)$_R$ factor. Most simply, the LRM can arise, e.g., from SO(10) GUT breaking directly to
SU(2)$_L \times$SU(2)$_R \times$U(1)$_{B-L}$ which then breaks to the SM at the few to multi-TeV scale. A second set of GUT-based $\Zp$ models arise from
$E_6$~\cite{Robinett:1982tq,London:1986dk,Hewett:1988xc,Joglekar:2016yap} where most simply $E_6 \rightarrow$ SM$\times$U(1)$_\psi \times$U(1)$_\chi \rightarrow $ SM$
\times$U(1)$_\theta$, where a new U(1)$_\theta$ gauge group factor is predicted. Note here $\theta$ labels the remaining linear combination of U(1)$_\psi$-U(1)$_\chi$ that remains
unbroken to lower energies (in comparison to the GUT scale).  A common set of features of this GUT-based model class include their possesing generation-universal couplings
of the $\Zp$ to the SM fermions, their charges commuting with those of the SM so that, e.g., $u_L$ and $d_L$ have the same $\Zp$ coupling and the resonances themselves are usually
narrow, reflecting electroweak strength or weaker couplings with width to mass ratios $\Gamma/M < 0.01-0.03$. In particular, the GUT origin of these models
implies that this class of $\Zp$ can be used to simultaneously study all of the dileptonic channels: $e^+e^-,~\mu^+\mu^+$ as well as $\tau^+\tau^-$ together with the dijets and boosted
$t\bar t$ channels as well.

With this much information potentially available from the observation of a given $\Zp$ in multiple channels one may try to distinguishable it from others of similar type given
sufficient statistics and
well-controlled systematics. In addition to relative cross section measurements,  e.g., that of dijets and/or $t\bar t$ compared to dileptons, the cleanliness of the dilepton channel itself
can allow us to obtain additional information.  Since the leptons can be signed, their angular distribution allows us to determine their forward-backward asymmetry, $A_{FB}$, (defined in the
dilepton center of mass frame) which depends upon the quark and lepton couplings of the $\Zp$ in a different manner than does the dilepton production cross section and, hence, will
differ from model to model. Note that theoretically the scattering angle is usually defined as the one between the outgoing negatively charged lepton and the incoming valence quark direction which is {\it usually} also the direction of the boost of the center of mass frame as seen in the lab frame. However, sometimes this condition does not hold and the anti-quark direction is
instead that of the boost and this must be corrected for statistically in Monte Carlo, but not an an event-by-event basis. This observable is discussed more fully below along with some of its
alternative definitions. A second possibility~\cite{delAguila:1993ym} is to make use of the fact
that the rapidity distributions of the $u\bar u$ and $d\bar d$ PDFs are somewhat different. Since various $\Zp$ will generally couple differently to the  $u$ and $d$ quarks the
rapidity distributions of the dilepton final state will probe these coupling variations. This possibility can be probed by forming the rapidity ratio, $r_y$, which is the ratio of the number of
central dilepton pairs to that at larger rapidities; this too will be discussed in further detail below.

Returning to our discussion of these specific GUT-inspired models, we note that in the LRM with the assumption of left- and right-handed gauge couplings, i.e., $\kappa=g_R/g_L=1$,
all of the various interactions of the $\Zp$ with the SM fields are completely fixed. However, in the $E_6$ model case, the
single new mixing parameter, $\theta$, controls the couplings of the $\Zp$ to the various SM particles; four particular choices for the value of this parameter correspond to the more
specific model cases discussed here and are denoted as $\psi$, $\chi$, $\eta$ and I. As in the SM, the $\Zp$ in GUT models generally couple to all the familiar quarks and leptons and
thus can easily populate simultaneously the various fermionic 2-body final states listed above at various predictable rates. The measurement of these rates (as well as other associated
observables) can be then used to discriminate among the various $\Zp$ possibilities after discovery as will be discussed further below. Note that the decay rate for $\Zp$ into the
$W^+W^-$ final state in GUT frameworks is highly dependent on the details of the model building assumptions within a specific scenario and especially upon the detailed nature
of spontaneous symmetry breaking as manifested by the amount of mixing (if it occurs at all) between the $\Zp$ and SM $\Z$; the $\Zp$ coupling to $W^+W^-$ in U(1) extensions is
always controlled solely by the amount of this gauge boson mixing.

The $\Zp$ of the Sequential Standard Model~\cite{Altarelli:1989ff} is often included within this GUT class of models although it is not a true gauge theory in the conventional sense.
However, it has been used very frequently for many years as a standard candle by experimenters since it conveniently posits the existence of heavier copies of the usual SM gauge bosons
with exactly the same couplings as do the gauge bosons in the SM; this provides a useful yardstick with which one can make comparisons easily.

Alternative models of electroweak symmetry breaking, including the topcolor assisted technicolor scenarios, can also frequently lead to $\Zp$-like states~\cite{Hill:1994hp}
that can produce resonance signatures. The greatest difference of such theories from the GUT-type $\Zp$ model class lies in their having generation-dependent couplings of
potentially QCD strength. (The color-octet versions of such states in this model class are called colorons.) This implies that the corresponding resonance will likely not be narrow and
will preferentially couple, by construction, to the
third generation, e.g., the highly boosted $t\bar t$ final state, thus proving another useful benchmark model for this channel. Similar new $\Zp$ states can also arise in Little Higgs
models~\cite{ArkaniHamed:2001nc} which can also have preferential decays to third generation states.

Occasionally the expected properties of a new $\Zp$ models are completely data-driven.  A $\Zp$ with an unusual flavor-dependent coupling structure has been suggested as
a (partially complete) UV model to explain the apparent anomaly seen in semileptonic $b\rightarrow sl^+l^-$ decays~\cite{Aaij:2014ora,Aaij:2017vbb}. In effective field theory language,
a new interaction of the form $\sim \bar bP_Ls \bar \mu P_L \mu$ of proper strength can provide a reasonable fit to these experimental observations~\cite{Bifani:2018zmi} which can
be the result of the exchange of a
very heavy $\Zp$ potentially accessible to high energy colliders ~\cite{Allanach:2017bta}. This $\Zp$, in the weak basis, couples only to the third generation quark doublet and to the
muon lepton doublet so that it will have a suppressed production cross section at hadron colliders. Such a $\Zp$ could be observed in both the dimuon and ditop channels.

Models of composite quarks and leptons offer another path wherein new resonances are predicted. Excited quarks ~\cite{Baur:1987ga,Baur:1989kv}, $Q^*$, are spin-1/2, color triplet
objects which carry the same SM quantum numbers as do the SM quarks. Here one imagines that the usual quarks have some type of internal structure and are held very tightly together
by some new BSM force; these constituents when excited in some way yield the more massive states we would then observe as $Q^*$. There is, as of yet, no fundamental, UV-complete
model encompassing this idea so that this framework is purely phenomenological. The SM quarks couple to these excited states via a magnetic dipole-like interaction together with an associated gauge boson such as the gluon or the SM $W,\Z$ or $\gamma$.  This interaction is suppressed by a large `compositeness scale', $\Lambda$, since it corresponds to a dim-6 operator, and the relative coupling strengths to the different gauge bosons are partially controlled by a set of essentially free parameters, $f_i$. Excited quarks can be singly produced in
the $gq$ channel to which they will also dominantly decay due to the presence of the strong coupling constant, yielding the dijet signature of interest to us here although decays into, e.g,
the $q\gamma$ channel are also of some interest. It is useful to have a benchmark model with dijet decays which take place in the $gq$ channel (as opposed to a $\Zp$ which can
only populate the $q\bar q$ dijet channel) with which to compare and contrast. The angular distributions of the 2 jets in the dijet decay, which will require significant statistics to
determine, can provide
us information about the spin of the original resonance and the nature of its couplings to the decay products~\cite{Harris:2011bh,Boelaert:2009jm,Chivukula:2014pma,Chivukula:2017nvl}.

Spin-2 graviton resonances occur in extra-dimensional scenarios that attempt to address the hierarchy problem, in particular, in the case of the warped extra dimensional model of
Randall and Sundrum (RS)~\cite{Randall:1999ee}. In such setups, the SM gauge fields and fermions are generally allowed to propagate in the 5-D
bulk~\cite{Pomarol:1999ad,Davoudiasl:1999tf,Grossman:1999ra,Davoudiasl:2000wi,Gherghetta:2000qt} whereas electroweak symmetry breaking occurs on or near the TeV/SM brane
via the usual Higgs mechanism. This approach simultaneously helps to address the SM fermion mass hierarchy by the localization properties of the SM fermions in the bulk.
One finds that, due to the shape of their 5-D wavefunctions, the Kaluza-Klein excitations of the familiar graviton, $G_{RS}$~\cite{Davoudiasl:1999jd} will dominantly decay into
objects localized near to where SM symmetry breaking occurs, i.e., Higgs boson pairs and $t\bar t$ as well as to the longitudinal components of the massive SM gauge bosons, e.g.,
$W^+_L W^-_L$, all with relatively fixed branching factions with only some small allowed variations. Thus $G_{RS}\rightarrow
W^+W^-$ in the RS framework provides an excellent benchmark model for the study of resonant $W$-pairs which are also quite highly boosted. If these $W$'s decay hadronically,
given this high boost, this final state may also (appear to) populate the resonant dijet channel. One notes that apart from the $G_{RS}$ mass scale itself, essentially the only other
free parameter in this RS model setup (wherein the lighter fermions are essentially decoupled from the graviton resonances), is frequently denoted by $c=k/\bar M_{Pl}$, which simply
controls the overall coupling strength to all of the various SM particles.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation setup}
\label{sec:simulation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Monte-Carlo production}
\label{subsec:mcprod}

Monte Carlo~(MC) simulated event samples are used to simulate the response of the detector to signal and backgrounds. Signals are generated with \py~\cite{Sjostrand:2014zea} version v8.201 with the \pdf{NNPDF2\!.\!3NNLO} PDF set~\cite{Ball:2014uwa} using the leading order cross-section from the generator with no k-factor. The SM backgrounds are Drell-Yan, di-jet (QCD), top pairs (\ttbar), $VV$ and \vj\ where $V=W/Z$,  were generated using \MGAMC~\cite{Alwall:2014hca} v2.5.5 at leading order only with the \pdf{NNPDF3\!.\!0NLO}~\cite{Ball:2014uwa} PDF set in bins of $\hht$. A k-factor of 2 is applied to all the background processes to account for higher order corrections.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Detector parameterizations}
\label{subsec:detparam}

As part of the FCC study group, the \delphes\ software package~\cite{deFavereau:2013fsa} is used to emulate the response a detector.
For the FCC-hh and HE-LHC the basic performances can be summarised as follow: The muon momentum resolution is assumed to be $\sigma(p)/p \approx 6\%~(10\%)$ at $\pt= 20~(5)$ TeV and $|\eta| \approx 0.$ for FCC-hh (HE-LHC) scenario.
\CH{maybe we need a table to compare basic performances of FCC and HE-LHC like~\ref{tab:delphes_comp}, but in any case it needs further description}.

\begin{table}[!htb]\centering
\begin{tabular}{|c|c|c|}
\hline
 & FCC-hh & HE-LHC \\
\hline
 muon & & \\
\hline
 electron  & & \\
\hline
 tau  & & \\
\hline
 jets & & \\
\hline
  b-jets  & & \\
\hline
\end{tabular}
\caption{Comparison of basic detector performances for FCC-hh and HE-LHC.}
\label{tab:delphes_comp}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Treatment of the Monte-Carlo samples}
\label{subsec:mctreat}
%\subsection{Tag Rate Function}
%\label{subsec:trf}
The modelling of the backgrounds in the high tagging regimes is a challenging task. The requirement of $b$ tagging in some MC samples can drastically reduce the available statistics.
This shortage of events that pass the $b$-tagging cut in the signal regime, in conjunction with the large cross section of some of the backgrounds can lead to very spiky templates.
To overcome this problem the tag rate function (TRF) method is used. By using the TRF method, no event is cut based on its $b$-tagging count, but instead all the events are weighted.
This weight can be interpreted as the probability of the given event to contain the desired number of $b$ jets.
To achieve this, the tagging efficiency (a function of $\eta$, $\pt$ and true jet flavour) was
used to calculate the event weight based on the kinematics and flavour of the jets found in each event. More details on the method can be found in Appendix~\ref{sec:app:trf}.
Despite the fact that very large amount of Monte-Carlo statistic has been simulated in bins of $\hht$
and the usage of TRF to save events, there are still large statistical fluctuations from high weight events.
In order to reduce this effect, and when large fluctuations are observed, the background spectrum is fitted. Further details on the procedure are given in Appendix~\ref{sec:app:bgfit}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Statistical analysis}
Hypothesis testing is performed using a modified frequentist method based on a profile likelihood that takes into account the systematic uncertainties as nuisance parameters that are fitted to the expected Monte-Carlo. The full shape information is used, which help from the sidebands to reduce the effect of systematics uncertainties in the signal region. The test statistic $q_\mu$ is defined as the profile log-likelihood ratio: $q_\mu = -2\ln({\cal L}(\mu,\hat{\hat{\theta}}_\mu)/{\cal L}(\hat{\mu},\hat{\theta}))$, where $\hat{\mu}$ and $\hat{\theta}$ are the values of the parameters that maximise the likelihood function (with the constraint $0\leq \hat{\mu} \leq \mu$), and $\hat{\hat{\theta}}_\mu$ are the values of the nuisance parameters that maximise the likelihood function for a given value of $\mu$. In the absence of any significant deviation from the background expectation, $q_\mu$ is used in the CL$_\text{s}$ method~\cite{Junk:1999kv,Read:2002hq} to set an upper limit on the signal production cross-section times branching ratio at the 95\% CL. For a given signal scenario, values of the production cross-section (parameterised by $\mu$) yielding CL$_\text{s} < 0.05$, where CL$_\text{s}$ is computed using the asymptotic approximation~\cite{Cowan:2010js}, are excluded at 95\% CL. For a $5\sigma$ discovery, the quantity 1-CL$_\text{b}$ must be smaller than $2.87 \cdot 10^{-7}$~\cite{Junk:1999kv} and is again computed using the asymptotic approximation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Di-lepton channels}
\label{sec:lep}

The decay products of heavy resonances are in the multi-TeV regime and the capability to reconstruct their momentum imposes stringent requirement on the detector design. In particular, reconstructing the track curvature of multi-TeV muons requires excellent position resolution and a large lever arm. In this section, the expected sensitivity is presented for a \Zpll\ (where $\ell=e,\mu$) and \Zptata\ separately.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The \texorpdfstring{\ee}{ee} and \texorpdfstring{\mumu}{mumu} final states}
\label{sec:lepee}

Events are required to contain two isolated opposite sign leptons with $\pt > 1$\,TeV, $|\eta|$<4 and an invariant mass $\mll > 2.5$\,TeV.
Figure~\ref{figure:leptonicresonances:ll} (top) shows the invariant mass for a 30\,TeV \ZpSSM\ signal for the $ee$ (left) and $\mu\mu$ (right) channels. The mass resolution is better for the $ee$ channel, as expected from calorimeter constant term that dominates the resolution at high \pt. Event yields can be found in Appendix~\ref{app:yields}.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.35\columnwidth]{Fig/mzp_sel0_nostack_log_ee.eps}
  \includegraphics[width=0.35\columnwidth]{Fig/mzp_sel0_nostack_log_mm.eps}
  \includegraphics[width=0.35\columnwidth]{Fig/lim_Zprime_ll_fcc_v02_allxs.eps}
  \includegraphics[width=0.35\columnwidth]{Fig/DiscoveryPotential_ll_comb_rootStyle.eps}
  \caption{Top: Invariant mass for a 30\,TeV signal after full event selection for $ee$ channel (left) and $\mu\mu$ channel (right).
  Bottom: Limit versus mass for the combined di-lepton ($ee$, $\mu\mu$) channel (left) and luminosity for a $5\sigma$ discovery (right) comparing ee, $\mu\mu$ and combined channels.}
  \label{figure:leptonicresonances:ll}
\end{figure}

The di-lepton invariant mass spectrum is used as the discriminant and a 50\% normalisation uncertainty on the background normalisation is assumed.
Figure~\ref{figure:leptonicresonances:ll} (bottom left) shows the 95\% CL exclusion limit obtained with 30\,ab$^{-1}$ of data combining ee and $\mu\mu$ channels. Figure~\ref{figure:leptonicresonances:ll} (bottom right) shows the integrated luminosity required to reach a $5\sigma$ discovery as a function of the mass of the heavy resonance. The \Zpee\ and \Zpmumu\ channel display very similar performance due to the low background rates. We conclude therefore that the reference detector design features near to optimal performance for searches involving high \pt\ muon final states. Combining $ee$ and $\mu\mu$ channels, masses up to 42\,TeV can be excluded or discovered. The limits for $ee$ and $\mu\mu$ alone can be found in Appendix~\ref{app:anamore}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The \texorpdfstring{\tautau}{tautau} final state}
\label{sec:leptautau}

At current LHC~\cite{Khachatryan:2016qkc}, the most sensitive channel to search for high mass di-$\tau$ resonances is when both $\tau$ leptons decays hadronically. This is the focus of the analysis presented in this section.
The event selection requires two jets with $p_{T} > 0.5$\,TeV and $|\eta|<2.5$, both identified as $\tau$'s. To ensure no overlap between the $\ell=e,\mu$ and $\tau$ final states, jets containing an electron or a muon with $\pt > 100$\,GeV are vetoed. Finally, requirements of $\Delta \phi(\tau_1, \tau_2)> 2$ and $2.5<\Delta R(\tau_1, \tau_2)<4$ are applied to suppress multi-jet background.
Furthermore, mass dependent cuts are applied to maximise the signal significance and are summarised in Table~\ref{tab:leptonicresonances:tautau}. Several proxies for the true resonance mass have been tested, such as the invariant mass of the two $\tau$'s, with and without correction for the missing energy, but the transverse mass~\footnote{the transverse mass is defined as $m_{T}  =  \sqrt{2\ptZp*\met*(1-cos\Delta\phi(\Zp,\met))} $} provided the best sensitivity and is therefore used to estimate the sensitivity.

Figure~\ref{figure:leptonicresonances:tautau} shows the di-$\tau$ transverse mass (left) for a 10\,TeV \ZpSSM. The 95\% CL exclusion limits for 30\,ab$^{-1}$ of data (middle) and the required integrated luminosity versus mass of the resonance to reach a $5\sigma$ discovery (right). Heavy resonance decaying to $\tau$ leptons reconstructed in the hadronic decay mode are more challenging given the overwhelming multi-jet background, but masses up to 18\,TeV could be probed.

\begin{table}[htbp]
   \centering
\begin{tabular}{|l|l|c|r|}
  \hline
  \hline
   $\Zp$ mass [TeV] &  $\Delta \phi(\tau_1, \tau_2)$&  $\Delta R(\tau_1, \tau_2)$ & $\met$\\
  \hline
  $4-8$ & > 2.4 & > 2.5 and < 3.5 & > 400 GeV\\
  $10$ & > 2.4 & > 2.7 and < 4 & > 300 GeV\\
  $12-14$ & > 2.6 & > 2.7 and < 4 & > 300 GeV\\
  $16-18$ & > 2.7 & > 2.7 and < 4 & > 300 GeV\\
  $>18$ & > 2.8 & > 3 and < 4 & > 300 GeV\\
  \hline
  \hline
  \end{tabular}
  \caption{List of mass dependent cuts optimised to maximise the sensitivity for the \Zptata\ search.}
  \label{tab:leptonicresonances:tautau}
\end{table}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.32\columnwidth]{Fig/mt_finalsel_nostack_log.eps}
  \includegraphics[width=0.32\columnwidth]{Fig/lim_Zprime_tautau_fcc_v02_allxs.eps}
  \includegraphics[width=0.32\columnwidth]{Fig/DiscoveryPotential_tautau_rootStyle.eps}
  \caption{Di-$\tau$ transverse mass (left), limit versus mass for the di-tau channel (center) and luminosity for a $5\sigma$ discovery (right). }
  \label{figure:leptonicresonances:tautau}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Hadronic final states}
\label{sec:hadronic}

Heavy resonances decaying hadronically also imposes stringent requirement on the detector design. For instance,  precise jet energy resolution requires full longitudinal shower containment and highly boosted top quarks and $W$ bosons decay into highly collimated jets that need to  be disentangled from standard QCD jets by studying their substructure. High discrimination power and sensitivity for these searches at such extreme energies, requires excellent granularity both in the tracking detectors and calorimeters.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Multi-Variate object tagging}
\label{subsec:mvatagger}

% from physics paper
An important ingredient of the hadronic searches is the identification of heavy boosted top quarks and $W$ bosons. Two object level taggers using Boosted Decision Trees (BDT) were developed to discriminate $W$ and top jets against the light jet flavours treated as background.
Top and $W$ taggers were optimised using jets with a transverse boost of $\pt=$10 TeV. At these extreme energies, $W$ and top jets have a characteristic angular size $R=0.01-0.02$, i.e smaller than the typical electromagnetic and hadronic calorimeter cells. Following the approach described in~\cite{Larkoski:2015yqa}, we exploit the superior track angular resolution and reconstruct jets from tracks only using the anti-$k_T$ algorithm with a parameter R=0.2, but also larger values are used to increase the discrimination power of the BDT. The missing neutral energy is corrected for by rescaling the track 4-momenta by the factor $\ptSub{,trk}/\ptSub{,PF}$, where $\ptSub{,trk}$ is the track Jet \pt\ and $\ptSub{,PF}$ is the Particle-Flow (PF) Jet \pt. In what follows, we will simply refer to ``track jets'' as the jet collection that includes the aforementioned rescaling.
\newline
The boosted top tagger is built from jet substructure observables: the soft-dropped jet mass~\cite{Larkoski:2014wba} (\mSD) and N-subjettiness~\cite{Thaler:2010tr} variables $\tau_{1,2,3}$ and their ratios $\tau_{2}/\tau_{1}$ ($\tau_{21}$) and $\tau_{3}/\tau_{2}$ ($\tau_{32}$). The $W$-jet tagger also uses an ``isolation-like'' variable that exploits the absence of high \pt\ final state-radiation (FSR) in the vicinity of the $W$ decay products. We call these variables $E_{F}(n,\alpha)$ and define them as:

%\begin{equation}
%E_{F}(n,\alpha) =  \frac{\sum \limits_{\frac{n-1}{5}\alpha < \Delta R(k,jet)< \frac{n}{5}\alpha} \ptSup{(k)}}{\sum \limits_{\Delta R(k,jet)< \alpha} \ptSup{(k)}}
%\end{equation}

\begin{equation}
E_{F}(n,\alpha) =  \sum \limits_{\frac{n-1}{5}\alpha < \Delta R(k,jet)< \frac{n}{5}\alpha} \ptSup{(k)} \Biggm/ \sum \limits_{\Delta R(k,jet)< \alpha} \ptSup{(k)}
\end{equation}
with $k$ running over the jet constituents and $\alpha=0.05$. We construct 5 variables $E_{F}(n,\alpha)$ with $n=[1,2,3,4,5]$ and use them as input to the BDT.
The $W$ tagging performance has significantly better performance due to the good discrimination power of the energy-flow variables. We choose the working points for the analyses presented later with a top and $W$ tagging efficiencies of $\epsilon_S^{\text{top}}=60\%$ and $\epsilon_S^{\text{W}}=90\%$ corresponding respectively to a background rejection of $\epsilon_B^{\text{top}}=\epsilon_B^{\text{W}}=90\%$. These working points corresponds to a cut at 0.15 on the BDT value for both taggers. Independent samples have been generated for training the BDT and there is no overlap with the one used in the analysis.  More details can be found in Appendix~\ref{sec:app:mva}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The \jj\ final state}
\label{sec:hadjj}

Jets are clustered using particle-flow candidates with the anti-$k_T$~\cite{Cacciari:2008gp} algorithm and  parameter R=0.4. We require at least two very energetic jets with $\pt$>3\,TeV and $|\eta|<3$. As di-jet events from signal will tend to be more central than for background, the rapidity difference between the two leading jets $\Delta(\eta_1, \eta_2)$ is required to be smaller than 1.5. The di-jet invariant mass for the \qjj\ signal with a mass of 40\,TeV and the QCD contribution after the full event selection is shown on Figure~\ref{figure:hadronicresonances:jj} (left). The middle Figure shows the 95\% CL exclusion limit obtained with 30\,ab$^{-1}$ of data and the right Figure shows the integrated luminosity required to reach a $5\sigma$ discovery as a function of the Q$^*$ mass.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.32\columnwidth]{Fig/Mj1j2_pf04_sel1_nostack_log.eps}
  \includegraphics[width=0.32\columnwidth]{Fig/lim_Qstar_jj_fcc_v02.eps}
 \includegraphics[width=0.32\columnwidth]{Fig/DiscoveryPotential_jj_rootStyle.eps}
   \caption{Invariant mass distribution of the two selected jets for a 40\,TeV signal (left), limit versus mass (middle) and $5\sigma$ discovery reach.\MS{need editing / style}}
  \label{figure:hadronicresonances:jj}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The \texorpdfstring{\ttbar}{tt} final state}
\label{sec:hadtt}

As track jets are found to better resolve the jet sub-structure compared to particle-flow jets, the selection for the \rsg\ and \Zptt\ searches is using track jets. As no lepton veto is applied, there is also some acceptance for leptonic decays and the sensitivity to semi-leptonic $WW$ or \ttbar\ decays is enhanced by adding the $\metvec$ vector to the closest jet 4-momentum (among the two leading jets).
We require two jets with a $\pt$>3\,TeV and $|\eta|<3$ and $\Delta(\eta_1,\eta_2)<2.4$. Both jets must either be $W$ or top tagged (section~\ref{subsec:mvatagger}) by requiring multivariate tagger score larger than 0.15. In addition, high-\pt\ jets must be $b$-tagged for the \Zptt\ analysis. Finally, to further reject QCD, we require for both jets the soft-dropped mass to be larger than 40\,GeV. Figure~\ref{figure:hadronicresonances:tt} (right) shows the di-top invariant mass distribution after the final cuts for a 20\,TeV signal. Thanks to the BDT discriminant, the largest background contribution is top pair production. The middle Figure shows the 95\% CL exclusion limit obtained with 30\,ab$^{-1}$ of data and the right Figure shows the integrated luminosity required to reach a $5\sigma$ discovery as a function of the Z$^\prime$ mass.
Further developments to improve the mass resolution could be considered to improve the sensitivity, but already with such wide spectrum, 
\begin{figure}[!htb]
  \centering
   \includegraphics[width=0.32\columnwidth]{Fig/Mj1j2_pf08_MetCorr_fit_sel8_nostack_log.eps}
   \includegraphics[width=0.32\columnwidth]{Fig/lim_Zprime_tt_fcc_v02.eps}
   \includegraphics[width=0.32\columnwidth]{Fig/DiscoveryPotential_tt_SSM_TC2_tagger_TRFbtag_rootStyle.eps}
  \caption{Invariant mass distribution of the two leading jets for the pre-selection (left) and full selection (right) for a 40\,TeV signal for the \qjj\ (left), and a 20\,TeV signal for \rsg\ (center) and \Zptt\ (right) analyses.\MS{need editing / style}}
  \label{figure:hadronicresonances:tt}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The \texorpdfstring{\ww}{ww} final state}
\label{sec:hadww}

The expected exclusion limits at 95\% CL are shown in Figures~\ref{figure:hadronicresonances:ww}. For the \qjj\ masses up to 40\,TeV could be discovered with \intlumifcc. Reconstructing Heavy resonances decaying to $WW$ and \ttbar\ is more challenging and requires the use of novel approaches to boosted object tagging to reduce the backgrounds. The reach for \Zptt\ (in TC2 models) and \rsg\ is 24\,TeV and 22\,TeV respectively and it is possible to discover a $Z^{\prime}_{SSM} \rightarrow \ttbar$ up to $m_{Z}=18$\,TeV.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.32\columnwidth]{Fig/Mj1j2_pf08_fit_sel4_nostack_log.eps}
  \includegraphics[width=0.32\columnwidth]{Fig/lim_RSGraviton_ww_fcc_v02.eps}
  \includegraphics[width=0.32\columnwidth]{Fig/DiscoveryPotential_ww_tagger_rootStyle.eps}
  \caption{Invariant mass distribution of the two leading jets for the pre-selection (left) and full selection (right) for a 40\,TeV signal for the \qjj\ (left), and a 20\,TeV signal for \rsg\ (center) and \Zptt\ (right) analyses.\MS{need editing / style}}
  \label{figure:hadronicresonances:ww}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Comparison with 27 TeV HE-LHC}
\label{sec:ana27tev}
The search for heavy resonances have also been study at the 27\,TeV HE-LHC. The analyses strategies remains identical to what was already presented, to the only difference that thresholds have been re-optimised to accommodate for the lower center of mass energy. The changes can be summarised as follows:
\begin{itemize}
\item \Zpee\ or \Zpmumu : leton $p_T$ lowered from 1 to 0.5\,TeV
%\item \Zpmumu\ f.a. and t-channel LQ : leton $p_T$ lowered from  1.2 TeV to 0.75 TeV
\item \Zptata\ : mass dependent cuts changed as shown in Table~\ref{tab:leptonicresonances:tautau27}
\item \rsg, \Zptt, \qjj\ : jet $p_T$ lowered from 3 to 1\,TeV
\end{itemize}

The results are summarised in Table~\ref{tab:27vs100}, together with a comparison to FCC-hh.

\begin{table}[htbp]
   \centering
\begin{tabular}{|l|l|c|r|}
  \hline
  \hline
   $\Zp$ mass [TeV] &  $\Delta \phi(\tau_1, \tau_2)$&  $\Delta R(\tau_1, \tau_2)$ & $\met$\\
  \hline
   $2$ & > 2.4 & > 2.4 and < 3.9 & > 80 GeV\\
   $4$ & > 2.4 & > 2.7 and < 4.4 & > 80 GeV\\
   $6$ & > 2.4 & > 2.9 and < 4.4 & > 80 GeV\\
   $8$ & > 2.6 & > 2.9 and < 4.6 & > 80 GeV\\
  $10$ & > 2.8 & > 2.9 and < 4.1 & > 60 GeV\\
  $12$ & > 2.8 & > 3.0 and < 3.6 & > 60 GeV\\
  $14$ & > 3.0 & > 3.0 and < 3.3 & > 60 GeV\\
  \hline
  \hline
  \end{tabular}
  \caption{Mass dependent cuts optimised to maximise the sensitivity for the \Zptata\ resonance search.}
  \label{tab:leptonicresonances:tautau27}
\end{table}

\begin{table}[!htb]\centering
\begin{tabular}{|l|c|c|c|}
\hline
\hline
analysis   & \multicolumn{3}{c|}{HE-LHC (FCC-hh)} \\
           & Limit [TeV] & Disco [TeV]   & Disco [TeV] \\
           &  15 (30) ab$^{-1}$ & 1 (2.5) ab$^{-1}$ & 15 (30) ab$^{-1}$ \\
\hline
\Zpll\ (l=e,$\mu$) & 13 (40) & 10 (33) & 13 (43) \\
\Zptata       &  6 (14) &  3 (12) &  6 (18) \\
%\Zpmumu\ f.a. &  4 (25) &  - (10) &  2 (19) \\
\Zptt         & 10 (28) &  6 (16) &  8 (23) \\
\rsg          &  8 (28) &  5 (15) &  7 (22) \\
\qjj          & 14 (43) & 10 (36) & 12 (40) \\
\hline
\hline
\end{tabular}
\caption{Limit and discovery reach at 5 $\sigma$ comparison between HE-LHC and FCC-hh.}
\label{tab:27vs100}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Characterisation of a Z' discovery}
\label{sec:zprimedisc}

\newcommand*{\sqrtslhc}{\ensuremath{\sqrt{s}=\text{14 TeV}}}
\newcommand*{\sqrtshelhc}{\ensuremath{\sqrt{s}=\text{27 TeV}}}
\renewcommand*{\intlumihelhc}{\ensuremath{\mathcal{L}=15\text{ ab}^{-1}}}
\newcommand*{\intlumihllhc}{\ensuremath{\mathcal{L}=3\text{ ab}^{-1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Context of the study}
\paragraph*{Context of the study}
It is still legitimate to assume that a heavy resonance could be seen at the end of HL-HLC. If that is the case a new collider with higher energy
in the \com is needed to study its properties as too few events will be available at \sqrtslhc. In this section we present the discrimination potential between six $Z'$ models of a High Energy LHC (HE-LHC) with an assumed \com energy of 27\,TeV and an integrated luminosity of \intlumihelhc. Under the assumption that these $Z'$'s decay only to SM particles, we show that there are sufficient observables to perform this model differentiation in most cases.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Bounds from HL-LHC}
\paragraph*{Bounds from HL-LHC}
As a starting point it is needed to estimate what are, for $\sqrt s=14$\,TeV, the typical exclusion/discovery reaches for standard reference $Z'$ models assuming \intlumihllhc\ employing only the $e^+e^-$ and $\mu^+\mu^-$ channels. To address this and the other questions below we will use the same set of $Z'$ models as employed
in Ref.~\cite{Rizzo:2014xma} and mostly in Ref.~\cite{Han:2013mra}, both of which we will refer to frequently. We employ the MMHT2014 NNLO PDF set~\cite{Harland-Lang:2014zoa}
throughout with an appropriate constant $K$-factor (=1.27) to account for higher order QCD corrections. The production cross section times leptonic branching fraction is shown in Figure~\ref{fig:pheno:toy} (left) for these models at \sqrtslhc\ in the narrow width approximation (NWA). It has been and will be assumed here that these $Z'$ states only decay to SM particles.

\begin{figure}[htbp]
  \centering
    \includegraphics[trim={2cm 2cm 2cm 2cm},clip,width=0.45\columnwidth]{Fig/zp14tev-ref.pdf}
    \includegraphics[trim={2cm 2cm 2cm 2cm},clip,width=0.45\columnwidth]{Fig/zp27tev-ref.pdf}
    \caption{Left: $\sigma B_l$ in the NWA for the $Z'$ production at the $\sqrt s=14$\,TeV LHC as functions of the $Z'$ mass: SSM(red), LRM (blue), $\psi$(green), $\chi$(magenta),
$\eta$(cyan), I(yellow). (Right) $\sigma B_l$ of $Z'$ in models described in (left) at $\sqrt s=27$\,TeV.}
\label{fig:pheno:toy}
\end{figure}

Using the present ATLAS and CMS results at 13\,TeV,~\cite{Aaboud:2017buh} and~\cite{Sirunyan:2018exx}, it is straightforward to estimate by extrapolation the
exclusion reach at \sqrtslhc\ using the combined $ee+\mu\mu$ final states. This is given in the first column of Table~\ref{tab:pheno:spec}. For discovery, only the $ee$ channel is used due to poor $\mu\mu$-pair invariant mass resolution near $M_{Z'}=6$\,TeV. Estimates of the $3\sigma$ evidence and $5\sigma$
discovery limits are also given in Table~\ref{tab:pheno:spec}. Based on these results, we will assume in our study below that we are dealing with a $Z'$ of mass 6\,TeV. Figure~\ref{fig:pheno:toy} (right) shows the NWA cross sections for the same set of models but now at \sqrtshelhc\ with \intlumihelhc. We note that very large statistical samples will be available for the case of $M_{Z'}=6$\,TeV
for each dilepton channel.

%
\begin{table}
\centering
\begin{tabular}{|l|c|c|c|} \hline\hline
  Model &   95$\%$ \cl  &  $3\sigma$  & $5\sigma$   \\
\hline
SSM    &     6.62     &  6.09    &  5.62     \\
LRM    &   6.39     & 5.85     & 5.39  \\
$\psi$    &  6.10   & 5.55   & 5.07  \\
$\chi$   &  6.22    & 5.68    & 5.26   \\
$\eta$   &  6.15     &  5.59  &  5.16   \\
~I        & 5.98   &  5.45   &  5.05  \\
\hline\hline
\end{tabular}
\caption{ Mass reach for several $Z'$ models at \sqrtslhc\ with \intlumihllhc. }
\label{tab:pheno:spec}
\end{table}
%
\paragraph*{Definition of the discriminating variables}
\label{par:vardef}

The various $Z'$ models can be disentangled with the help of 3 inclusive observables: the production cross section times leptonic branching fraction $\sigma B_l$, the forward-backward asymmetry $A_{FB}$ and the rapidity ratio $r_y$. The variable $A_{FB}$ can be seen as an estimate of the charge asymmetry
\begin{equation}
A_{FB} = A_C =  \frac{\sigma(\Delta|y| > 0) - \sigma(\Delta|y| < 0)}{\sigma(\Delta|y| > 0) + \sigma(\Delta|y| < 0)},
\end{equation}
where $\Delta|y| = |y_l| - |y_{\bar{l}}|$. It has been checked that this definition is equivalent to defining
\begin{equation}
A_{FB} = \frac{\sigma_F - \sigma_B}{\sigma_F + \sigma_B},
\end{equation}
with $\sigma_F = \sigma (cos\theta^{*}_{cs})>0$ and $\sigma_B = \sigma (cos\theta^{*}_{cs})<0$ where $\theta^*_{cs}$ is the Collins-Soper frame angle. The variable $r_y$ is defined as the ratio of central over forward events:
\begin{equation}
r_y = \frac{\sigma(|y_{Z'}| < y_1)}{\sigma(y_1 < |y_{Z'}| <y_2)},
\end{equation}
where $y_1=0.5$ and $y_2=2.5$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{Model discrimination}

The model discrimination presented in this section has been performed assuming the HE-LHC detector parametrisation~\cite{hlhelhc_web} in \delphes~\cite{deFavereau:2013fsa}. In such a detector, muons at $\eta \approx 0$ are assumed to be reconstructed with a resolution $\sigma(p)/p \approx 7\%$ for $\pt~=~3$\,TeV.

\subparagraph*{Leptonic final states}
\label{par:lepana}

The potential for discriminating various $Z'$ models is first investigated using the leptonic $ee$ and $\mu\mu$ final states only. The signal samples for the 6 models and the Drell-Yan backgrounds have been generated with \py~\cite{Sjostrand:2014zea} including the interference between the signal and background. The $Z'$ decays assume lepton flavour universality. For a description of the event selection and a discussion of the discovery potential in leptonic final states for the list of $Z'$ models being discussed here, the reader should refer to Section~\ref{sec:lep}. We simply point out here that with \intlumihelhc\,, all $Z'$ models with $m_{Z'}~\lesssim~10$\,TeV can be excluded at \sqrtshelhc.

Figure~\ref{fig:ana:res} (left) shows the correlated predictions for the $A_{FB}$ and the rapidity ratio $r_y$ observables defined previously for these six models given the above assumptions. Although the interference with the SM background was included in the simulation, its effect is unimportant due to the narrowness of the mass window around the resonance that was employed. Furthermore, the influence of the background uncertainty on the results has been found to have little to no impact on the model discrimination potential. Therefore the displayed errors on $A_{FB}$ and $r_y$ are of statistical origin only. The results show that apart from a possible near degeneracy in models $\psi$ and $\eta$, a reasonable $Z'$ model separation can indeed be achieved.

Using a profile likelihood technique, the signal strength $\mu$, or equivalently, $\sigma B_l$, can be fitted together with its corresponding error using the the di-lepton invariant mass shape. The quantity $\sigma B_l$ and its total estimated uncertainty is shown in Figure~\ref{fig:ana:res} (center) as a function of the integrated luminosity. The $\sigma B_l$ measurement seems to be able to resolve the degeneracy between the $\psi$ and $\eta$ models with \intlumihelhc. It should be noted however that since the cross-section can easily be modified by an overall rescaling of the couplings, further handles will be needed for a convincing discrimination.

\begin{figure}[!htb]
  \centering
   \includegraphics[width=0.30\columnwidth]{Fig/ry_afb_200GeV_eta2p5_interf.eps}
   \includegraphics[width=0.30\columnwidth]{Fig/sigma_vs_lumi.eps}
   \includegraphics[width=0.30\columnwidth]{Fig/Zp_branching_6TeV_15ab.eps}
  \caption{Left: Scatter plot of $r_y$ versus $A_{FB}$ with 200~GeV and mass window. The full interference is included. Center: Fitted signal cross-section together with its corresponding error versus integrated luminosity. Right: Fitted cross-section of the three hadronic analyses. Statistical and full uncertainties are shown on each point.}
  \label{fig:ana:res}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsubsection{Hadronic analysis}
\subparagraph*{Hadronic final states}
\label{par:hadana}

Model discrimination can be improved by including an analysis involving three $Z'$ addition hadronic final states: $t\bar{t}$, $b\bar{b}$ and $q\bar{q}$, where $q=u,d,c,s$. The sample production and event selection for the $t\bar{t}$, $q\bar{q}$ final states has been described to some extent in Section~\ref{sec:hadronic}. We simply remind the reader that the analysis involves requiring the presence of two central high $\pt$ jets. In order to ensure complete orthogonality between the various final states, jets are required to be tagged as follows. In the $Z' \rightarrow t\bar{t}$ analysis both jets should be \emph{top-tagged}. For the $Z' \rightarrow b\bar{b}$ final state both jets are required to be \emph{b-tagged} and we veto events containing at least one top-tagged jet. Finally, in the $Z' \rightarrow q\bar{q}$ analysis, we veto events that contain at least one b-tagged or top-tagged jet.

Figure~\ref{fig:ana:res} (right) summarises the discrimination potential in terms of fitted cross-section of the different models considering the three aforementionned hadronic decays, $t\bar{t}$,  $b\bar{b}$ and $q\bar{q}$. An good overall discrimination among the various models can be achieved using all possible final states. For example, the SSM and $\psi$ models, which have very close predictions for $r_y$ and $A_{FB}$, have measurably different fractions of $t\bar{t}$ or $b\bar{b}$ final states. We note however that the degeneracy between $\eta$ and $\psi$ can only be partially resolved resolved at $\approx~1~\sigma$ by exploiting the difference in $t\bar{t}$ yield.

\paragraph*{Conclusion}
In this section we studied the discrimination potential of six $Z'$ models at High Energy LHC (HE-LHC) with an assumed \com energy of 27\,TeV and an integrated luminosity of \intlumihelhc. The exercise has been performed assuming the evidence of an excess observed at \sqrtslhc\ at a mass $m_{Z'}\approx~6$\,TeV. Overall it was found that it is possible to distinguish among most models. Finally, it should be noted that further studies, perhaps employing 3-body decay modes or associated Z' production will be clearly in case of discovery to further characterize the resonance properties.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Flavour anomaly inspired Z' models sensitivity at the FCC-hh}
\label{sec:zprimeflav}

\MS{completely missing this section}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:conc}

This note presents preliminary studies of a search for $\Zp$
bosons decaying into two electrons or muons in the FCC context. The expected number
of signal and background events have been estimated from simulated truth level information
after applying smearing functions to mimic the FCC detector response.
Using a cut-based analysis and assuming simplistic systematic uncertainties, $\ZpSSM$
masses below 40\,TeV can be excluded at 95$\%$ C.L.
using 30 $\afb{}$ of data.

\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multivariate object tagger}%appendix
\label{sec:app:mva}
The training samples are built from jets that do not contain leptons, for example from sem-leptonic b-decays.
The BDT parameters used are the following : NTrees=600, MaxDepth=4, AdaBoost, AdaBoostBeta=0.15, SeparationType=GiniIndex, nCuts=100, PruneMethod=NoPruning.
As an example, the $E_{F}(n=1,\alpha=0.05)$ observable is shown in Figures~\ref{fig:TMVA_final_result} left.
The evolution of the light jet efficiency versus the $W$ and top tagging efficiencies for both taggers is shown in Figure~\ref{fig:TMVA_final_result} right.
Cross-checks have been performed to further validate the procedure. First, by removing highly correlated variables, such as \CH{@David Add example}, it showed the same performances \CH{@ David, then why don't we remove the correlated variables?}.
As a second test, the BDT response has been tested for different signal masses to understand how the mass dependence could affect the analyses. For the cut used in the analysis (BDT score greater than 0.15), the shape of the BDT is not dramatically changing the signal efficiency.% and it removes 90\% and 95\% of light jets for thad and Whad taggers respectively.

%The input variables are defined to be the same as the ones used to perform cut-based associated analysis (\Zptt\ and \rsg\ in section~\ref{subsec:hadreso}) and the idea is to fully exploit these informations thanks to the BDT.
The input variables used to train the BDT, ordered by training weight, can be found in table~\ref{tab:TMVA_summary}.

\begin{table}[!htb]\centering
\begin{tabular}{| l | c | l | c |}
\hline
  \multicolumn{2}{|c|}{$W$ tagger}  & \multicolumn{2}{c|}{top tagger} \\
  \hline
 variable & weight & variable & weight \\
%\hline
%\hline
%Sample signal     & pp\_$RSG$\_$WW$\_20TeV (1M) & pp\_$Z'_{TC2}$\_$tt$\_20TeV (1M) \\
%Sample background & pp\_$jj$\_lo ($j$ = $u$, $d$, $c$, $s$ or $b$) (920k) & pp\_$jj$\_lo ($j$ = $u$, $d$, $c$, $s$ or $b$) (920k)               \\
%\hline
%Train stat. signal (jets)     & 1117500 & 1432192 \\
%Train stat. background (jets) & 1578098 & 1578098 \\
\hline
% rename them
 $\tau_3$ (track jet, R=0.2)      & 0.12      & $\tau_1$ (track jet, R=0.2) & 0.21  \\
 $\mSD$  (track jet, R=0.2)      & 0.11      & $\mSD$  (track jet, R=0.2) & 0.17 \\
 $\tau_{31}$  (track jet, R=0.2) & 0.10     & $\tau_{31}$  (track jet, R=0.2)  & 0.11 \\
 $E_{F}(n=5,\alpha=0.05)$                               & 0.09     &  $\tau_2$ (track jet, R=0.2) & 0.10 \\
 $E_{F}(n=4,\alpha=0.05)$                               & 0.09     & $\tau_3$ (track jet, R=0.2) & 0.09 \\
 $E_{F}(n=1,\alpha=0.05)$                               & 0.08     & $\mSD$  (track jet, R=0.8)& 0.09 \\
 $E_{F}(n=2,\alpha=0.05)$                               & 0.07     &  $\mSD$  (track jet, R=0.4) & 0.09 \\
 $E_{F}(n=3,\alpha=0.05)$                               & 0.06     & $\tau_{32}$  (track jet, R=0.2) & 0.08 \\
 $\tau_{21}$  (track jet, R=0.2)& 0.06   & $\tau_{21}$  (track jet, R=0.2) & 0.06 \\
 $\mSD$  (track jet, R=0.8) & 0.06 &  &\\
 $\mSD$  (track jet, R=0.4) & 0.06 & & \\
 $\tau_1$ (track jet, R=0.2) & 0.05      &  &\\
 $\tau_2$ (track jet, R=0.2) & 0.04      &  &\\
 $\tau_{32}$  (track jet, R=0.2) & 0.02    &  &\\
\hline
\end{tabular}
\caption{Summary of the input variables to the BDT and their relative weight for both $W$ and top taggers.}
\label{tab:TMVA_summary}
\end{table}

\begin{figure}[!htbp]\centering
\includegraphics[width=0.3\textwidth]{Fig/TMVA/Jet1_Flow15_sel0_nostack_logx.eps}
\includegraphics[width=0.3\textwidth]{Fig/TMVA/Jet1_tau32_sel0_nostack_log.eps}
\includegraphics[width=0.3\textwidth,trim={0 0.5cm 0 0},clip]{Fig/TMVA/effQCD_vs_effWhadBlue_thadRed_log.pdf}
\caption{Left: Energy-flow ($E_{F}(1,0.05)$) observable for W and QCD jets. Middle: ($\tau_(3,2)$) observable for t and QCD jets. Right: Light jet rejection versus tagging efficiency for the $W$ tagger (blue) and top tagger (red) -> re-do on-going.}
\label{fig:TMVA_final_result}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tagging rate function}%appendix
\label{sec:app:trf}
Given a jet with $\eta$, $\pt$ and flavour $f$, its tagging probability can be noted as:
\begin{equation*}
	\varepsilon \left(f,|\eta|,\pt\right)
\end{equation*}
\newline
For a given event with $N$ jets, its probability of containing exactly one $b$-tag jet can be computed as:
\begin{equation*}
	P_{=1} = \sum\limits_{i=1}^N \left( \varepsilon_{i} \prod\limits_{i \neq j} \left( 1 - \varepsilon_{j} \right) \right)
\end{equation*}
\newline
In the same way, it can be used to compute the probability for inclusive $b$-tag selections:
\begin{align*}
	P_{=0} &= \prod\limits_{i=1}^N \left( 1 - \varepsilon_{j} \right) \\
	P_{\geq 1} &= 1 - P_{=0}
\end{align*}
\newline
It was verify that the TRF methods agrees well with the direct tagging.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background fit}%appendix
\label{sec:app:bgfit}
\begin{equation}
f(z)=p_1(1-z)^{p_2}z^{p_3}z^{p_{4}logz}
\end{equation}
where $z=m_{jj}/\sqrt{s}$. This fit is used in order to have a smooth shape for the backgrounds, while the normalisation is taken prior to the fit (see figure~\ref{fig:hadronicresonances_nofit}).

\begin{figure}[!htb]\centering
\includegraphics[width=0.45\columnwidth]{Fig/Zptt/Mj1j2_pf08_MetCorr_sel8_nostack_log.eps}
\includegraphics[width=0.45\columnwidth]{Fig/Zptt/Zptt_QCD_sel8_Mj1j2_pf08_MetCorr_fit.eps}
\caption{Invariant mass prior to fit.}
\label{fig:hadronicresonances_nofit}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Event yields}
\label{app:yields}
\begin{table}[htbp]
   \centering
\begin{tabular}{|l|r|r|}
  \hline
 & $ee$ & $\mu\mu$  \\
  \hline
  Drell-Yan & 206882.9 & 236597.9 \\
  \hline
  $\Zp$ 4\,TeV & 1421357.9    & 1598969.4 \\
  $\Zp$ 6\,TeV & 349922.4  & 393117.6\\
  $\Zp$ 8\,TeV &   115043.5 & 129698.7 \\
  $\Zp$ 10\,TeV &  45423.5 & 50873.3 \\
  $\Zp$ 20\,TeV &  1192.3 & 1411.5\\
  $\Zp$ 30\,TeV &  88.2 & 107.6\\
  $\Zp$ 40\,TeV &  11.7 & 14.1 \\
  $\Zp$ 50\,TeV &  3.2 & 3.7\\
  \hline
  \hline
\end{tabular}
  \caption{Expected number of events at FCC-hh for the \Zpee\ and \Zpmumu\ analysis after the full event selection for the SSM signal hypothesis and considering 30ab$^{-1}$ of integrated luminosity.}
  \label{tab:leptonicresonances:yieldsll}
\end{table}

\begin{table}[!htb]
   \centering
\begin{tabular}{|l|r|r|}
  \hline
  \hline
 & ee & $\mu\mu$  \\
  \hline
  Drell-Yan    & 162699 &  244802 \\
  \hline
  $\Zp$ 2\,TeV  & 737966 & 1040077 \\
  $\Zp$ 4\,TeV  &  30351 &   53020 \\
  $\Zp$ 6\,TeV  &   2694 &    5343 \\
  $\Zp$ 8\,TeV  &    380 &     781 \\
  $\Zp$ 10\,TeV &     76 &     151 \\
  $\Zp$ 12\,TeV &     23 &      41 \\
  $\Zp$ 14\,TeV &     10 &      17 \\
  \hline
  \hline
\end{tabular}
  \caption{Expected number of events at HE-LHC for the \Zpee\ and \Zpmumu\ analysis after the full event selection for the SSM signal hypothesis and considering 15ab$^{-1}$ of integrated luminosity.}
  \label{tab:leptonicresonances27:yieldsll}
\end{table}

\begin{table}[htbp]
   \centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
  \hline
  \hline
  & \multicolumn{2}{c|}{di-jet}  & \multicolumn{2}{c|}{$\ttbar$} & \multicolumn{2}{c|}{WW} \\
  \cline{2-7}

 & pre-sel & final-sel  & pre-sel & final-sel & pre-sel & final-sel\\
  \hline
  di-jet & 385555434 &  373661126 &  154855591 & 11439.8&  154856148 & 64484\\
  $\ttbar$ & - & - & 1114779 & 74193.6 &  1114779 & 3185\\
  di-bosons & - & - &  41820 &  17.1 &  41820 & 6092\\
  boson+jet & - & - & 1610472 & 264.1&  1610472 & 25377\\
  \hline
  total bkg  &  385555434& 373661126& 157622662 & 85914 & 154856148 & 99137\\
  \hline
  10\,TeV &  - & - &  101529 & 15601 &  47853 & 15745\\
  20\,TeV &   1253072 &  1239813& 7774 & 500.6 & 1282 & 578\\
  30\,TeV &  69922 &  67488 & 485.2 & 13.2 &  61.4 & 30.1 \\
  40\,TeV &  4589 &  4373 & - & - & - & -\\
  \hline
  \hline
\end{tabular}
  \caption{Expected number of events at FCC-hh for the di-jet, $\ttbar$ and WW analyses after pre and final selection.}
  \label{tab:hadronicresonances:yields}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%appendix
\section{More information on analyses}
\label{app:anamore}
\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.35\columnwidth]{Fig/lim_Zprime_ee_fcc_v02_allxs.eps}
  \includegraphics[width=0.35\columnwidth]{Fig/lim_Zprime_mumu_fcc_v02_allxs.eps}
  \caption{Limit versus mass for the ee channel (left) and $\mu\mu$ channel (right) for an integrated luminosity of $30$\,ab$^{-1}$. }
  \label{figure:leptonicresonances:resultseemumu}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%appendix
\section{Summary plots}
The discovery potential for resonances considered in FCC-hh scenario are summarised in Figure~\ref{figure:resonances100:summary}.
\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.45\columnwidth]{Fig/summaryDisco_onlyFCChh.pdf}
  \includegraphics[width=0.45\columnwidth]{Fig/summaryDisco_onlyHELHC.pdf}
    \includegraphics[width=0.45\columnwidth]{Fig/summaryDisco.pdf}

  \caption{Summary of a $5\sigma$ discovery reach as a function of the resonance mass for different luminosity scenario of FCC-hh and HE-LHC.}
  \label{figure:resonances100:summary}
\end{figure}

\acknowledgments

This is the most common positions for acknowledgments. A macro is
available to maintain the same layout and spelling of the heading.

\paragraph{Note added.} This is also a good position for notes added
after the paper has been written.

% The bibliography will probably be heavily edited during typesetting.
% We'll parse it and, using the arxiv number or the journal data, will
% query inspire, trying to verify the data (this will probalby spot
% eventual typos) and retrive the document DOI and eventual errata.
% We however suggest to always provide author, title and journal data:
% in short all the informations that clearly identify a document.

\bibliographystyle{JHEP}
\bibliography{paper}

% Please avoid comments such as "For a review'', "For some examples",
% "and references therein" or move them in the text. In general,
% please leave only references in the bibliography and move all
% accessory text in footnotes.

% Also, please have only one work for each \bibitem.%appendix

\end{document}