### 2019-02-22 paper_epjc.tex

%\RequirePackage{fix-cm}

\documentclass[pdftex,twocolumn,epjc3]{svjour3}          % twocolumn
\usepackage{commands}
\RequirePackage[T1]{fontenc}

\smartqed  % flush right qed marks, e.g. at end of proof

\RequirePackage{graphicx}
\RequirePackage{mathptmx}      % use Times fonts if available on your TeX system
\RequirePackage{flushend}
\RequirePackage[numbers,sort&compress]{natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}

\journalname{Eur. Phys. J. C}

\begin{document}

\title{\boldmath Heavy resonances at energy-frontier hadron colliders}

%\subtitle{Do you have a subtitle?\\ If so, write it here}

\author{David~Jamin\thanksref{e1,addr1,addr3}
        \and
        Clement~Helsens\thanksref{e2,addr1}
        \and
        Michelangelo~L.~Mangano\thanksref{e3,addr1}
        \and
        Thomas~G.~Rizzo\thanksref{e4,addr2}
        \and
        Michele~Selvaggi\thanksref{e5,addr1}
}

%\thankstext[$\star$]{t1}{Thanks to the title}
\thankstext{e1}{e-mail: david.jamin@cern.ch}
\thankstext{e2}{e-mail: clement.helsens@cern.ch}
\thankstext{e3}{e-mail: michelangelo.mangano@cern.ch}
\thankstext{e4}{e-mail: rizzo@slac.stanford.edu}
\thankstext{e5}{e-mail: michele.selvaggi@cern.ch}

\institute{CERN, CH-1211 Geneva 23, Switzerland\label{addr1}
          \and
          SLAC National Accelerator Laboratory 2575 Sand Hill Rd., Menlo Park, CA, 94025 USA\label{addr2}
          \and
          Academia Sinica, Institute of  Physics, Taipei, Taiwan\label{addr3}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
This paper explores the physics reach of the proton-proton Future Circular Collider (FCC-hh) and of the High-Energy LHC (HE-LHC) for searches of new particles produced in the $s$ channel and decaying to two high-energy leptons, jets (non-tops), tops or W/Z bosons. We discuss the expected discovery potential and exclusion limits for benchmark models predicting new massive particles that result in resonant structures in the invariant mass spectrum. We also present a detailed study of the HE-LHC potential to discriminate among different models, for a $\Zp$ that could be discovered by the end of High-Luminosity LHC (HL-LHC).
\end{abstract}

\input{paper_text}

\begin{thebibliography}{9}
\bibliography{paper}
\end{thebibliography}

\end{document}